---
title: "Indian Used Car Sales Analysis"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
---

```{r}
library(dplyr)
library(stringr)
library(tidyr)
library(glmnet)
library(lmtest)
library(corrplot)
```

## Reading and Cleaning the Data set

```{r}

cars_raw<- read.csv("used_cars_data.csv",stringsAsFactors = FALSE)
dim(cars_raw)
names(cars_raw)
str(cars_raw)
summary(cars_raw)
```

```{R}
#drops ID since its useless
#will also not consider New price since over 70% of NewPrice col is empty
cars<- cars_raw %>% select(-S.No.,-New_Price)
#converting the currency from INR Lakhs to USD
cars <- cars %>% mutate(Price = 1120*Price)

#Stripping the unit from these variables' values so they can be quantifiable 
cars<-cars  %>% mutate(Mileage = as.numeric(str_extract(Mileage, "[0-9.]+")))
cars<-cars  %>% mutate(Engine = as.numeric(str_extract(Engine, "[0-9.]+")))
cars<-cars  %>% mutate(Power = as.numeric(str_extract(Power,"[0-9.]+")))

#car_age is more usable than make year of model. 
current_year <- 2020
cars <- cars %>% mutate(Age = current_year - Year)
cars <- cars %>% select(-Year)
```


```{R}
num_cars <- cars %>% select_if(is.numeric)
corr_matrix<-cor(num_cars)
corrplot(corr_matrix, method="color", type="upper",addCoef.col = "black", tl.srt = 45)

```

```{R}
#make sure there are no more missing values
colSums(is.na(cars))
cars <- na.omit(cars)
```

## Regularization + Transformations

```{R}
hist(cars$Price, xlab="Price", main="Distribution of Car Prices")
hist(log(cars$Price), xlab="Price", main="Price Histogram after log")


#Since the plot is very rightly skewed, we should apply log to normalize it
```

```{R}
#Want to make sure that we do not include the variables name and location to prevent the creation of dummy variables to be created
cars_no_name_location <- cars %>% select(-Name,-Location)
ols_model <- lm(log(cars$Price)~., data=cars_no_name_location)
summary(ols_model)

library(car)
vif_values <- vif(ols_model)
vif_values
sort(vif_values, decreasing = TRUE)
#Engine displacement and engine horsepower carry almost the same information → strong multicollinearity.
```

# Checking for Assumptions

```{R}

plot(ols_model,which = 1)
plot(ols_model,which = 2)
plot(ols_model,which = 3)


bptest(ols_model)



```

# Residual vs fitted plot does not really show any strong curvature, this plot indicates that the relationship between log(Prices) and the predictors is reasonably Linear.

# Q-Q Plot indicates the central portion is normal. The tails however deviate noticeably due to several large residuals at the end. Due to the large sample size, I do not think that this violation is severe enough to invalidate the model.

# According to the B-P test, there is some heteroscedasticity present. Given that our sample size is large (over 5,000 observations), the BP test can be extremely sensitive. This makes me believe it should not be so severe to invalidate it for that reason. If we also look at the Residual vs fitted plot, we can see that there is only mild heteroscedasticity if not minor.

```{R}
ols_res <- rstandard(ols_model)
outliers <- which(abs(ols_res) > 3)

cars_clean <- cars_no_name_location[-outliers, ]
ols_clean <- lm(log(Price) ~ ., data = cars_clean)

plot(ols_clean, which = 1)
plot(ols_clean, which = 2)
bptest(ols_clean)
```

# Trying to get rid of the outliers to deal with the homoskedasticity seems to make it worst so we will just keep it the way it is

## Perform OLS and Lasso

```{R}
x <- model.matrix(log(Price)~., data=cars_no_name_location)[,-1]
y <- log(cars_no_name_location$Price)
#x <- model.matrix(Price~., data=cars_clean)[,-1]
#y <- cars_clean$Price

set.seed(1)
#Lasso:
cv_lasso <-cv.glmnet(x,y, alpha=1)
plot(cv_lasso)
title("Lasso Cross Validation", line = 2.5)
#Best Lambda
lasso_lambda <- cv_lasso$lambda.min
#coefficients at Lambda.min
lasso_coefs <- coef(cv_lasso, s= lasso_lambda)
as.matrix(lasso_coefs)



#Ridge:
cv_ridge <- cv.glmnet(x,y, alpha=0)
plot(cv_ridge)
title("Ridge Cross Validation", line = 2.5)
#Best Lambda
ridge_lambda <- cv_ridge$lambda.min
#coefficients at Lambda.min
ridge_coefs <- coef(cv_ridge, s= ridge_lambda)

```

```{R}
coef_df<- data.frame(variable = rownames(lasso_coefs), coefficient=as.numeric(lasso_coefs))
coef_df %>% filter (coefficient!=0) %>% arrange(desc(abs(coefficient)))
```

```{R}
set.seed(1)
#splitting data 30-70
train_idx <- sample(1:nrow(cars_no_name_location), 0.7 * nrow(cars_no_name_location))
train <- cars_no_name_location[train_idx, ]
test  <- cars_no_name_location[-train_idx, ]
# OLS model on training data
ols_train <- lm(log(Price) ~ ., data = train)
ols_pred <- predict(ols_train, newdata = test)

#Prepare matrices for glmnet
#glmnet needs X matrix and y vector:
x_train <- model.matrix(log(Price) ~ ., data = train)[, -1]
y_train <- log(train$Price)

x_test  <- model.matrix(log(Price) ~ ., data = test)[, -1]
y_test  <- log(test$Price)

#fit Lasso
cv_lasso <- cv.glmnet(x_train, y_train, alpha = 1)
lasso_lambda <- cv_lasso$lambda.min

lasso_model <- glmnet(x_train, y_train, alpha = 1, lambda = lasso_lambda)
lasso_pred <- predict(lasso_model, s = lasso_lambda, newx = x_test)

#fit Ridge
cv_ridge <- cv.glmnet(x_train, y_train, alpha = 0)
ridge_lambda <- cv_ridge$lambda.min

ridge_model <- glmnet(x_train, y_train, alpha = 0, lambda = ridge_lambda)
ridge_pred <- predict(ridge_model, s = ridge_lambda, newx = x_test)

library(ggplot2)

rmse_df <- data.frame(
           Model = c("OLS", "Lasso", "Ridge"),
           RMSE  = c(ols_rmse, lasso_rmse, ridge_rmse)
           )


ggplot(rmse_df, aes(x = Model, y = RMSE)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = round(RMSE, 4)), vjust = -0.3) +
  coord_cartesian(ylim = c(0.30, 0.32)) +
  labs(
    title = "Test RMSE by Model (Zoomed In)",
    x = "Model",
    y = "RMSE (log-price)"
  ) +
  theme_minimal()




#Compare RMSE for all 3 Models
rmse <- function(actual, predicted) {
        sqrt(mean((actual - predicted)^2))
        }

ols_rmse   <- rmse(y_test, ols_pred)
lasso_rmse <- rmse(y_test, lasso_pred)
ridge_rmse <- rmse(y_test, ridge_pred)

ols_rmse
lasso_rmse
ridge_rmse
```
## Conclusion:
